Test 1

### 2020-11-11 DATA622/Test1.rmd
More work on Test 1

```{r}
get_bootstrap_sample <- function(data, bootstrap_proportion = 0.6, sample_with_replacement = TRUE){
  n_bootstrap_observations <- round(nrow(data) * bootstrap_proportion, 0)
  data[sample(nrow(data), n_bootstrap_observations, replace = sample_with_replacement),]
}

@@

## Gathering the Model's Capacity to Learn

In order to evaluate this capacity, we will gather metrics on the algorithm's performance on the training set.

@@

```{r}
lr_model <- glm(label ~ ., data = training_df, family = "binomial")
summary(lr_model)
```
*The sample data are too small (`r nrow(training_df)`) to have any confidence in the Logistic Regression model, so I will be throwing this one out.  Bagging and LOOCV will not resolve this issue.*

```{r}
#training_lr_yhat <- get_lr_yhat(lr_model, training_df, "label") This does not make sense to use
capacity_to_learn <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
```
### Naive Bayes

@@

#### With Bagging

Now to try bagging the Naive Bayes model.  I will save all the models to a list so that they can be used to evaluate how well it would generalize latter in the document.

```{r}
nb_bag <- list()
for (i in 1:n_bags){
  bag_df <- get_bootstrap_sample(training_df)
  bag_df_without_label <- bag_df %>% select(-label)
  bag_df_label <- bag_df$label
  bag_nb_model <- naiveBayes(label ~ ., data = bag_df)
  training_bag_nb_yhat <- predict(bag_nb_model, bag_df_without_label)
  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_nb_yhat, paste("NB Bag", i))
  nb_bag[[i]] <- list(model = bag_nb_model, yhat=training_bag_nb_yhat, capacity_to_learn = bag_capacity_to_learn)
  if(exists("nb_bag_capacity_to_learn")){
    nb_bag_capacity_to_learn <- rbind(nb_bag_capacity_to_learn, bag_capacity_to_learn)
  } else {
    nb_bag_capacity_to_learn <- bag_capacity_to_learn
  }
}
```

Now to average the capacity to learn metrics:

```{r}
nb_bag_capacity_to_learn <- nb_bag_capacity_to_learn %>%
  select(-Algo) %>%
  colMeans(na.rm=TRUE) %>%
  t() %>%
  data.frame("Algo"=c("NB with Bagging"), .)
capacity_to_learn <- rbind(capacity_to_learn, nb_bag_capacity_to_learn)
```

#### LOOCV

Now Naive Bayes with LOOCV

```{r}
for (i in 1:nrow(training_df)){
  loocv_test <- training_df[i,]
  loocv_test_without_label <- loocv_test %>% select(-label)
  loocv_training_df <- training_df[-c(i),]
  loocv_nb_model <- naiveBayes(label ~ ., data = loocv_training_df)
  training_loocv_nb_yhat <- predict(loocv_nb_model, loocv_test_without_label)
  loocv_capacity_to_learn <- evaluate_algo(loocv_nb_model, training_loocv_nb_yhat, paste("NB LOOCV", i))
}
```

@@

@@

## Gathering the Model's Capacity to Generalize

```{r}
lr_yhat <- get_lr_yhat(lr_model, test_df, "label")
test_lr_cm <- confusionMatrix(table(test_df_label, lr_yhat))
capacity_to_generalize <- evaluate_algo(test_df_label, lr_yhat, "LR")
```

@@

@@

We have evaluated `r nrow(capacity_to_learn)` algorithm's capacity to learn and generalize.  The following tables summarize their performace:

**Table 1. Capacity to Learn**
```{r, echo=FALSE}
capacity_to_learn %>%  kable() %>%  kable_styling()
```

>>>>

```{r}
get_bootstrap_sample <- function(data, bootstrap_proportion = 0.6, sample_with_replacement = TRUE){
  n_bootstrap_observations <- round(nrow(data) * bootstrap_proportion, 0)
  return(data[sample(nrow(data), n_bootstrap_observations, replace = sample_with_replacement),])
}

@@

I will also need a way to aggregate/average the metrics

```{r}
average_metrics <- function(metrics, algo){
  avg_metrics <- metrics %>%
  select(-Algo) %>%
  colMeans(na.rm = TRUE) %>%
  t() %>%
  data.frame("Algo" = c(algo), .)
  return(avg_metrics)
}
```

## Baseline Models

First I will train the models following what was previously done in HW1 and collect the metrics on both the training and test data sets

@@

```{r}
lr_model <- glm(label ~ ., data = training_df, family = "binomial")
summary(lr_model)
```
#### Capacity to Learn

**The sample data are too small (`r nrow(training_df)`) to have any confidence in the Logistic Regression model, so I will be throwing this one out.  Bagging and LOOCV will not resolve this issue.**

```{r}
capacity_to_learn <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
```

#### Capacity to Generalize

**Again the sample data are too small to use to make meaningful inferences using Logistic Regression.  Bagging and LOOCV will not help and will be excluded from the metrics.**

```{r}
capacity_to_generalize <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
```

I will repeat the same process for Naive Bayes.

#### Capacity to Learn

@@

#### Capacity to Generalize

@@

Finally I will repeat the process for KNN where k=3 and k=5

#### Capacity to Learn

@@

#### Capacity to Generalize

@@

## Models With Bagging

Now to try bagging the Naive Bayes, and KNN models.  I will use the models trained on the bootstrap samples to predict the test data and collect metrics on the capacity to learn and generalize, which will be aggregated.

```{r}
nb_bag <- list()
for (i in 1:n_bags){
  bag_df <- get_bootstrap_sample(training_df)
  bag_df_without_label <- bag_df %>% select(-label)
  bag_df_label <- bag_df$label

  # Train the Naive Bayes Model
  bag_nb_model <- naiveBayes(label ~ ., data = bag_df)
  training_bag_nb_yhat <- predict(bag_nb_model, bag_df_without_label)
  ## Evaluate the capacity to learn
  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_nb_yhat, paste("NB Bag", i))
  if(exists("nb_bag_capacity_to_learn")){
    nb_bag_capacity_to_learn <- rbind(nb_bag_capacity_to_learn, bag_capacity_to_learn)
  } else {
    nb_bag_capacity_to_learn <- bag_capacity_to_learn
  }
  ## Evaluate the capacity to generalize
  bag_nb_test_yhat <- predict(bag_nb_model, test_df_without_label)
  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_nb_test_yhat, paste("NB Bag", i))
  if(exists("nb_bag_capacity_to_generalize")){
    nb_bag_capacity_to_generalize <- rbind(nb_bag_capacity_to_generalize, bag_capacity_to_generalize)
  } else {
    nb_bag_capacity_to_generalize <- bag_capacity_to_generalize
  }

  # Train the KNN3 Model
  ## Evaluate the capacity to learn
  training_bag_knn3_yhat <- knn(bag_df_without_label, bag_df_without_label, bag_df_label, k = 3)
  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_knn3_yhat, paste("KNN3 Bag", i))
  if(exists("knn3_bag_capacity_to_learn")){
    knn3_bag_capacity_to_learn <- rbind(knn3_bag_capacity_to_learn, bag_capacity_to_learn)
  } else {
    knn3_bag_capacity_to_learn <- bag_capacity_to_learn
  }
  ## Evaluate the capacity to generalize
  bag_knn3_test_yhat <- knn(bag_df_without_label, test_df_without_label, bag_df_label, k = 3)
  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_knn3_test_yhat, paste("KNN3 with Bagging", i))
  if(exists("knn3_bag_capacity_to_generalize")){
    knn3_bag_capacity_to_generalize <- rbind(knn3_bag_capacity_to_generalize, bag_capacity_to_generalize)
  } else {
    knn3_bag_capacity_to_generalize <- bag_capacity_to_generalize
  }

  # Train the KNN5 Model
  ## Evaluate the capacity to learn
  training_bag_knn5_yhat <- knn(bag_df_without_label, bag_df_without_label, bag_df_label, k = 5)
  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_knn5_yhat, paste("KNN5 with Bagging", i))
  if(exists("knn5_bag_capacity_to_learn")){
    knn5_bag_capacity_to_learn <- rbind(knn5_bag_capacity_to_learn, bag_capacity_to_learn)
  } else {
    knn5_bag_capacity_to_learn <- bag_capacity_to_learn
  }
  ## Evaluate the capacity to generalize
  bag_knn5_test_yhat <- knn(bag_df_without_label, test_df_without_label, bag_df_label, k = 5)
  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_knn5_test_yhat, paste("KNN5 with Bagging", i))
  if(exists("knn5_bag_capacity_to_generalize")){
    knn5_bag_capacity_to_generalize <- rbind(knn5_bag_capacity_to_generalize, bag_capacity_to_generalize)
  } else {
    knn5_bag_capacity_to_generalize <- bag_capacity_to_generalize
  }

}
```

Now to average the metrics:

```{r}
# Naive Bayes
capacity_to_learn <- rbind(capacity_to_learn, average_metrics(nb_bag_capacity_to_learn, "NB with Bagging"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(nb_bag_capacity_to_generalize, "NB with Bagging"))

# KNN3
capacity_to_learn <- rbind(capacity_to_learn, average_metrics(knn3_bag_capacity_to_learn, "KNN3 with Bagging"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn3_bag_capacity_to_generalize, "KNN3 with Bagging"))

# KNN5
capacity_to_learn <- rbind(capacity_to_learn, average_metrics(knn5_bag_capacity_to_learn, "KNN5 with Bagging"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn5_bag_capacity_to_generalize, "KNN5 with Bagging"))
```

## Models With LOOCV

Now to examine how the models preform using LOOCV

```{r}
training_loocv_nb_yhat <- c()
training_loocv_knn3_yhat <- c()
training_loocv_knn5_yhat <- c()
loocv_test_label <- c()

for (i in 1:nrow(training_df)){
  # Leave One Out
  loocv_test <- training_df[i,]
  loocv_test_without_label <- loocv_test %>% select(-label)
  loocv_test_label <- c(loocv_test_label, loocv_test$label)
  loocv_training_df <- training_df[-c(i),]
  loocv_training_df_without_label <- loocv_training_df %>% select(-label)
  loocv_training_df_label <- loocv_training_df$label

  # Train the Naive Bayes Model
  loocv_nb_model <- naiveBayes(label ~ ., data = loocv_training_df)
  ## Evaluate the capacity to learn
  ### Note: This will be done once we have gone over the whole training data set
  training_loocv_nb_yhat <- c(training_loocv_nb_yhat, predict(loocv_nb_model, loocv_test_without_label))
  ## Evaluate the capacity to generalize
  loocv_nb_test_yhat <- predict(loocv_nb_model, test_df_without_label)
  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_nb_test_yhat, paste("NB with LOOCV", i))
  if(exists("nb_loocv_capacity_to_generalize")){
    nb_loocv_capacity_to_generalize <- rbind(nb_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
  } else {
    nb_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
  }

  # Train the KNN3 Model
  ## Evaluate the capacity to learn
  ### Note: This will be done once we have gone over the whole training data set
  training_loocv_knn3_yhat <- c(training_loocv_knn3_yhat, knn(loocv_training_df_without_label, loocv_test_without_label, loocv_training_df_label, k = 3))
  ## Evaluate the capacity to generalize
  loocv_knn3_test_yhat <- knn(loocv_training_df_without_label, test_df_without_label, loocv_training_df_label, k = 3)
  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_knn3_test_yhat, paste("KNN3 with LOOCV", i))
  if(exists("knn3_loocv_capacity_to_generalize")){
    knn3_loocv_capacity_to_generalize <- rbind(knn3_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
  } else {
    knn3_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
  }

  # Train the KNN5 Model
  ## Evaluate the capacity to learn
  ### Note: This will be done once we have gone over the whole training data set
  training_loocv_knn5_yhat <- c(training_loocv_knn5_yhat, knn(loocv_training_df_without_label, loocv_test_without_label, loocv_training_df_label, k = 5))
  ## Evaluate the capacity to generalize
  loocv_knn5_test_yhat <- knn(loocv_training_df_without_label, test_df_without_label, loocv_training_df_label, k = 5)
  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_knn5_test_yhat, paste("KNN5 with LOOCV", i))
  if(exists("knn5_loocv_capacity_to_generalize")){
    knn5_loocv_capacity_to_generalize <- rbind(knn5_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
  } else {
    knn5_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
  }

}
```

Now to average the metrics:

```{r}
# Naive Bayes
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_nb_yhat, "NB with LOOCV"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(nb_loocv_capacity_to_generalize, "NB with LOOCV"))

# KNN3
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_knn3_yhat, "KNN3 with LOOCV"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn3_loocv_capacity_to_generalize, "KNN3 with LOOCV"))

# KNN5
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_knn5_yhat, "KNN5 with LOOCV"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn5_loocv_capacity_to_generalize, "KNN5 with LOOCV"))
```

@@

We have evaluated `r nrow(capacity_to_learn)` algorithm's capacity to learn and generalize.  The following tables summarize their performance:

**Table 1. Capacity to Learn**
```{r, echo=FALSE}
capacity_to_learn %>% arrange(Algo) %>% kable() %>%  kable_styling()
```

**Table 2. Capacity to Generalize**
```{r, echo=FALSE}
capacity_to_generalize %>% arrange(Algo) %>% kable() %>% kable_styling()
```

### Discussion

```{r}
ggplot() + geom_density(aes(x=ACCURACY, color="Bagging"), data=nb_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="LOOCV"), data=nb_loocv_capacity_to_generalize) + geom_vline(xintercept = 1/3) + ggtitle("Naive Bayes Capacity to Generalize") + scale_color_brewer(palette = "Set1")
```

```{r}
ggplot() + geom_density(aes(x=ACCURACY), data=knn3_bag_capacity_to_generalize) + geom_vline(xintercept = 1/2)
```

```{r}
ggplot() + geom_density(aes(x=ACCURACY), data=knn5_bag_capacity_to_generalize) + geom_vline(xintercept = 2/3)
```
